---
title: "Data Wrangling for Capstone"
author: "Sitruks"
date: "May 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggmap)
library(naniar)
```

### Capstone Summary
Prediction model (win) - school proximity to host site

#### Capstone Data Sourcing
Fortunately, since the source of most data comes from [Kaggle](https://www.kaggle.com/) the data required for this project is clean. Much of the data has to be joined together to run the stats models.  
  
One bit of information that was not provided was the specific coordinates (latitude and longitude) of home and tournament venues. The venue locations can be found on [Wikipedia](https://en.wikipedia.org/wiki/List_of_NCAA_Men%27s_Division_I_Basketball_Tournament_venues), but the coordinates are still required. To pull coordinates, the **ggmap** package will need to be installed. Ggmap leverages google's API to get this information. 
  
This data contains old arena locations (which is redundant for humans but shouldn't affect google) and is missing a column for state data and will require some cleaning to improve google api query accuracy. What is nice about this data set is that it also lists the year and round that was played, which may be very useful for matching. Once cleaned, the geocodes can be pulled and merged with "GameCities.csv". "Teams.csv", which does not require any cleaning, can run ggmap at any time to get the pertinent home venue data.  

#### Questions
For this problem, how would you recommend compiling the data set? 
* At work, I often work with relational databases. Creating relationships in the data usually alleviates the need to merge all sources.  
    + Can you create relationships in R?  
  
Help with using mutate_geocode():  
* how can i reduce the errors  
  
How to replace/set values as "NA" not just a character named "NA"  
* I found the package [naniar](https://cran.r-project.org/web/packages/naniar/vignettes/replace-with-na.html)  
    + have you ever used this?  
    + do you know of a better way?  
  
What is the best path to add new files to github?  
* directly to github?  
* from Rstudio?  
  
Is there a "ctrl + z" equivalent you know of in R?  
* The use case would be for after assigning a varible that overwrites your data  
    + i.e. x <- 4 %>% x <- 4345  
  
#### Data Wrangling

```{r}
#TourneyVenues data clean
TourneyVenues <- read_csv("Data/TourneyVenues.csv")

Venues_Clean <- TourneyVenues %>% 
  mutate(State = if_else(is.na(Arena), City, "NA")) %>% 
  replace_with_na(list(State = "NA")) %>% 
  fill(State) %>% 
  filter(!is.na(Arena)) %>% 
  unite(City, State, Arena, col = "Address", sep = ", ")
```
#advice for the following
Venues_Clean %>% mutate_geocode(Address)

```{r}
#optional extra cleaning
Venues_Clean %>% 
    separate_rows(`Opening rounds`) %>% 
    separate_rows(Regionals) %>% 
    separate_rows(`Final Four`)
```