---
title: "Exercise 2"
author: "Kurtis"
date: "May 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)

```

### Data Wrangling Exercise 2: 


#### Task 0
Load the data in RStudio  

```{r} 
df <- tbl_df(read_csv("C:/Users/v-kursan/Downloads/Capstone_DS/Wrangle/2_Basic_Data_Manipulation/titanic_original.csv"))
```

#### Task 1
Port of Embarkation
```{r}

df <- df %>% replace_na(list(embarked = "S"))
# check below  
df %>% group_by(embarked) %>% summarise(n = n())

```

#### Task 2
Age  
```{r}
#convert double to integer then find mean
df <- df %>% mutate(age = as.integer(age))
#find mean
df %>% summarise(mean(age, na.rm = TRUE))
#replace NA's
df <- df %>% replace_na(list(age = 30))
#check below
df %>% group_by(age) %>% summarise(n = n())
```

#### Task 3
Lifeboat  
```{r}
#replace na first
df %>% replace_na(list(boat = "None"))
#distribution

df %>% group_by(boat) %>% summarise(n = n()) %>% spread(boat, n)

```

#### Task 4
Cabin 

* it makes sense to fill in the missing cabin number values to potentially make summarise functions easier to run. Also to better distinguish what NA means (i.e. missing, unknown, no room)
* here, it is hard to tell what a missing value means, but i will assume "None"

```{r}
df %>% mutate(has_cabin_number = if_else(cabin = is.NA), 0, 1)

```