---
title: "Capstone Statistical Analysis"
author: "Sitruks"
date: "September 8th, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

Packages <- c("readxl", "tidyverse", "tidytext", "stringr", "ggmap", "naniar", "data.table", "readr")
lapply(Packages, library, character.only = TRUE)
```


```{r}

df11 <- read_csv('Data/df11.csv')
```

## NCAA Basketball Tournament Prediction Model: Home Field Advantage

### **Explore Data**

In order to make sense of the data, I first explored it (InitialVisualization_Capstone.R).

Immediately, I was reminded of the discrepancies in tournament play over time, namely the total field change from 32 to 64 in 1985. One team was added to the mix in 2001. In the 2011 season 4 teams were added -- solidifying the current (2018) 68-team tournament. As such, properly filtering out data pre-XXXX will be an important item to consider. 

In addition, accurate team locations (latitude and longitude) is essential to proving my hypothesis. No such information is available, so I will need to write a script to pull the coordinates. The ggmap package fulfils this need. 

Also there is a wealth of regular season data, the model should be first built on the tournament data before expanding.

The venue data that was sourced includes specific stadiums unique to ncaa women's teams, and needs to be filtered 

### **Apply Statistics**


```{r, include=FALSE}
ncaa_table <- data.table(df11)
e1 <- ncaa_table %>%
  select(Season, Seeding, Home_Team_Match, TeamID) %>% 
  unique() %>%
  filter(Season %in% (2010:2017)) %>%
  filter(Seeding == 1) %>% 
  group_by(Home_Team_Match) %>% 
  tally() %>% 
  top_n (n = 10) %>% 
  arrange(desc(n)) 

e2 <- ncaa_table %>%
  select(Season, Tourney_State, Home_Team_Match, TeamID) %>% 
  unique() %>%
  filter(Season %in% (2010:2017)) %>%
  group_by(Home_Team_Match) %>% 
  tally() %>% 
  top_n (n = 10) %>% 
  arrange(desc(n)) 

e3 <- ncaa_table %>%
  select(Season, TeamID.Type, Home_Team_Match, TeamID) %>% 
  unique() %>%
  filter(Season %in% (2010:2017)) %>%
  filter(TeamID.Type == "WTeamID") %>%
  group_by(Home_Team_Match) %>% 
  tally() %>% 
  top_n (n = 10) %>% 
  arrange(desc(n)) 

```

#### **stats 1**
```{r}

e1 %>% 
ggplot(aes(x = reorder(Home_Team_Match, n), y = n,label = n)) + geom_text(color="white", size=2) +
    geom_point(stat = 'identity', fill = 'darkblue',size = 4) + geom_segment(aes(y=0,x = reorder(Home_Team_Match, n), yend = n,xend =reorder(Home_Team_Match, n) ),color = 'black') +
    labs(x = '', y = 'No 1 seeds', title = 'No. 1 Seeds from 2010 - 2017') +
    coord_flip()
```

#### **stats 2**
```{r}

e2 %>% 
ggplot(aes(x = reorder(Home_Team_Match, n), y = n,label = n)) + geom_text(color="white", size=2) +
    geom_point(stat = 'identity', fill = 'darkblue',size = 4) + geom_segment(aes(y=0,x = reorder(Home_Team_Match, n), yend = n,xend =reorder(Home_Team_Match, n) ),color = 'black') +
    labs(x = '', y = '', title = 'Tournament Site Locations Since 2010') +
    coord_flip()
```

#### **stats 3**
```{r}

e3 %>% 
ggplot(aes(x = reorder(Home_Team_Match, n), y = n,label = n)) + geom_text(color="white", size=2) +
    geom_point(stat = 'identity', fill = 'darkblue',size = 4) + geom_segment(aes(y=0,x = reorder(Home_Team_Match, n), yend = n,xend =reorder(Home_Team_Match, n) ),color = 'black') +
    labs(x = '', y = '', title = 'Tournament Wins Since 2010') +
    coord_flip()
```



### **Interesting Findings**

Kansas has received the most #1 seeds since 2010 with 5. The next closest teams have each received 3 (North Carolina, Kentucky, and Duke).

Over that same time frame, Kansas is tied for third in frequency of hosting tournament games.

Interestingly enough, Kanasas has won the most games in the NCAA tournament.

Although it is too early to tell, this gives credence to the theory that distance can be a predictor of winning in the tournament.


### **Actions**
The ggmap package, though accurate, incurred an error best described as a time-out between calls. As such, a loop will need to be developed to manageably pull this information. For this case, a while loop should suffice.

The other thing that became apparent was a need to compute distances between related coordinates. The geosphere package is a good solution for this. Further testing may be required on this package.
